---
title: "Lab 01"
author: "Jiapeng Guo"
date: "May 6, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
data(mtcars)
```

### Arithmetic

In the first code block, use R to perform the following arithmetic operations. 

- $13+5$
- $\frac{7}{3}$
- $14^3$

```{r}
# Note that you can run an individual code chunk by clicking on the green arrow over here -->
# Instead of knitting the entire document, you can work on little bits of code first!

# Write your code below
#(a)
13 + 5
#(b)
7 / 3
#(c)
14^3

```

***

### Variable Assignments

In the second code block, use R, to complete the following variable assignments.

- Assign $6$ to $x$. Assign $-5$ to $y$. Evaluate the sum of $x$ and $y$.
- You can assign *from* variables just as easily as you can assign *to* variables. Reassign $x$ so that it is equal to $y$. Reevaluate the sum of $x$ and $y$.
- Reassign $y$ so that it has the value of $8$. In the previous part, we reassigned $x$ so that it has the same value as $y$. Does changing the value of $y$ in this part affect the value of $x$? 

```{r}
# Write your code below
# (a)
x = 6
y = 5
x + y
#(b)
x = y
x + y
#(c)
y = 8

# it does not affect the value of x

```

***

### Data Types

In the third code block, use R to explore different data types.

- Create three variables `myName`, `myPets`, and `myState`. Assign them your name, the number of pets your family has, and whether you are from Michigan (`TRUE` if you are from Michigan, `FALSE` otherwise). Note, you will need to wrap your name in double quotes.
- Check the type of each variable using the `class()` function

```{r}
# Write your code below
#(a)
myName = 'Jiapeng'
myPets = 0
myState = FALSE

#(b)
class(myName)
class(myPets)
class(myState)
```

***

### Basic Data Exploration

In the fourth code block, use R to explore the built-in data set 'mtcars'. This data contains aspects of automobile design and performance for a sample of cars from 1974.

- Using the `summary()` function, create numerical summaries for the entire data set.
- Use the `View()` function to look at the data. Calling `View()` should be done in the console below, not in the R Markdown file. No work is required for this task.
- Are more cars automatic (`am = 0`) or manual (`am = 1`)? Get the answer by using the `table()` function. Note: to reference a specific variable, use the format `data$variable`.
- What is the standard deviation of the variable fuel efficiency (`mpg`)? Use the function `sd()`.
- What is the highest horsepower to weight ratio (i.e. `hp:wt`)? Note: generate a new variable that represents this ratio, then use the `max()` function to find the maximum ratio.
- Use the `plot()` function to create a scatterplot of fuel efficiency (`mpg`) versus weight (`wt`). Use the inputs `xlab` and `ylab` to give the plot proper $x$ and $y$-labels.


```{r}
# The data set 'mtcars' has already been loaded into the workspace at the top of the file

# Write your code below
#(a)
summary(mtcars)

#(b)
table(mtcars$am)
# There are more cars that are automatic than cars that are manual

#(c)
sd(mtcars$mpg)

#(d)
mtcars$ratio = mtcars$hp / mtcars$wt
max(mtcars$ratio)

#(e)
plot(mtcars$mpg, mtcars$wt, xlab = 'fuel efficiency', ylab = 'weight')
```

